Variation in use of East Asian Late Paleolithic weapons: A study of tip cross-sectional area of stemmed points from Korea

### 2023-03-12 analysis/paper/paper.qmd
editing abstract

abstract: |
 The transition from the Early to Late Paleolithic in Korea is characterized by the introduction of blade technology, stemmed points, end scrapers, burins, denticulates, and finer grained materials. Stemmed points have been considered a representative tool that led this set of changes. In this study, we examine the possible role that stemmed points played during the technological transition, as well as throughout the Late Paleolithic period. Our main questions are: What were the best-fit ballistic probabilities for the stemmed points if they were hafted as weapon tips? How diverse were their likely uses? What are the temporal and spatial patterns of stemmed point use? We measured tip cross-sectional area (TCSA) to distinguish different likely use classes of projectile points, for example, as poisoned arrow tips or as stabbing spears. We analyzed TCSA with other variables, including raw materials, weight, radiocarbon dates and locations. Our results show that the stemmed points could have served as javelin tips and stabbing spear tips, with smaller numbers as dart tips and un-poisoned arrow tips. TCSA values are controlled mostly by size rather than raw material types. We found different TCSA ranges of stemmed points at different sites, which could indicate people used stemmed points in different ways depending on the local environment. Some sites show a wide range of TCSA values that represent multi-purpose usage of stemmed points. The temporal pattern of TCSA values is one of little change throughout the Late Paleolithic period, but points were predominantly produced before the Last Glacial Maximum (LGM). We observed that stemmed points were mostly located in certain ecoregions in Korea but no clear spatial pattern was confirmed. We conclude that stemmed points were multi-functional tools, with many likely designed for javelin and stabbing spear tips.

@@

# ggsave(here::here("analysis/figures/fig-tcsa-raw-materials.png"),
#        width = 4.3,
#        height = 2.7,
#        units = "in",
#        bg = "white")

>>>>

abstract: |
 The transition from the Early to Late Paleolithic in Korea is characterized by the introduction of blade technology, stemmed points, end scrapers, burins, denticulates, and higher proportions of finer grained materials. Stemmed points have been considered a representative tool that led this set of changes. In this study, we examine the possible role that stemmed points played during this technological transition, as well as throughout the Late Paleolithic period. Our main questions are: What were the best-fit ballistic probabilities for the stemmed points if they were hafted as weapon tips? How diverse were their likely uses? What are the temporal and spatial patterns of stemmed point use? We measured tip cross-sectional area (TCSA) to distinguish different likely use classes of projectile points, for example, as poisoned arrow tips or as stabbing spears. We analyzed TCSA with other variables, including raw materials, weight, radiocarbon dates and locations. Our results show that the stemmed points likely served as javelin tips and stabbing spear tips, with smaller numbers as dart tips and un-poisoned arrow tips. TCSA values were controlled mostly by size rather than raw material types. We found different TCSA ranges of stemmed points at different sites, which could indicate people used stemmed points in different ways depending on the local environment. Some sites show a wide range of TCSA values that represent multi-purpose usage of stemmed points. The temporal pattern of TCSA values is one of little change throughout the Late Paleolithic period, but points were predominantly produced before the Last Glacial Maximum (LGM). We observed that stemmed points were mostly located in certain ecoregions in Korea, but no clear spatial pattern was apparent We conclude that stemmed points were multi-functional tools, with many likely designed for use as javelin and stabbing spear tips.

@@

# ggsave(here::here("analysis/figures/fig-tcsa-raw-materials.png"),
#        width = 4.3,
#        height = 2.7,
#        units = "in",
#        bg = "white")
```
